---
title: "MAKE START DATA MS - R Notebook"
output: html_notebook
---

#INPUTS
Takes the datasets used in the workshop, and applies the learning that we need to start with:
- data below 100m only
- excluding Tromsøflaket (first cruise) data which has some analytical anomalies for the moment
- generate a random one subsample per VL dataset for first analysis step

## Libraries
```{r}
library(tidyverse)
library(oce)
```

# LOAD DATA

```{r}
#NB 2022-10-19 latest (all sample/Oslo) species cleaning applied so smaller spp dataset

env_orig <- read.csv(file.path(dataPath,"inputs/envDataOslo_2022-10-19.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputs/widedataOslo_2022-10-19.csv")) %>% as.data.frame

#left join and split to ensure ordered identically
JoinedDat<- left_join(otu_orig,env_orig, by=c("SampID"="SampID2"))

```


## REMOVE ALL STATIONS SHALLOWER THAN 100m 

```{r}

plot(JoinedDat$X.y, JoinedDat$Y)

# RETAIN ONLY SAMPLES BELOW 100m depth
JoinedDat$bathyPositive<-(JoinedDat$bathy*-1)
AfotJoinedDat <-subset(JoinedDat, bathyPositive>100)
AfotJoinedDat<-AfotJoinedDat%>%select(-c("bathyPositive"))

#check plot (compare to before)
plot(AfotJoinedDat$X.y, AfotJoinedDat$Y)

```


## REMOVE TROMSØFLAKET DATA (i.e. data from cruise 2006612)

```{r}
# remove 2006 data
T2006_filt<-filter(AfotJoinedDat, !str_detect(cruise_no,'2006612'))
T2006_check<-filter(T2006_filt, (str_detect(cruise_no,'2006612')))
```

## GENERATE MAIN DATASETS

```{r}
Afot_otu_sort<-T2006_filt %>% select(c(2:which(colnames(T2006_filt)=="Weberella_bursa")))#dataset 19-10-22 or after
Afot_env_sort<-T2006_filt %>% select(c(2,(which(colnames(T2006_filt)=="Weberella_bursa")+1):which(colnames(T2006_filt)=="sand")))  #dataset 19-10-22 or after
# LDensNoSB_otu_sort<-loDensNoSBJoinedDat %>% select(c(2:391))#dataset before 19-10-22
# LDensNoSB_env_sort<-loDensNoSBJoinedDat %>% select(c(2,397:532))  #dataset before 19-10-22
```


##### SAVE DATASETS
```{r}
# SHALLOW BELOW 100m dataset without Tromsøflaket = START DATASET
write.csv(Afot_otu_sort, file=file.path(dataPath,(paste0("inputsMS/Afot_otu_sort_", Sys.Date(),".csv"))))
write.csv(Afot_env_sort, file=file.path(dataPath,(paste0("inputsMS/Afot_env_sort_", Sys.Date(),".csv"))))
```





## GENERATE 1 per VL DATASET

### SUBSAMPLE
```{r}
# Subset samples for maneagbility - Take a random (sub)sample from each VL

subSamp1perVL <- T2006_filt %>% group_by(VL) %>% slice_sample(n=1)
length(unique(subSamp1perVL$VL))

plot(subSamp1perVL$X.y, subSamp1perVL$Y)
```


### SPLIT INTO OTU & ENV DATASETS

```{r}
Afot1pVL_otu_sort<-subSamp1perVL %>% ungroup %>% select(c(2:which(colnames(subSamp1perVL)=="Weberella_bursa")))
Afot1pVL_env_sort<-subSamp1perVL %>% select(c(2,(which(colnames(subSamp1perVL)=="Weberella_bursa")+1):which(colnames(subSamp1perVL)=="sand")))  

```


write the subsample to csv with sys.Date appended to keep track of which is the active sample that day
```{r}
write.csv(Afot1pVL_otu_sort, file=file.path(dataPath,(paste0("inputsMS/Afot1pVL_otu_sort_", Sys.Date(),".csv"))))
write.csv(Afot1pVL_env_sort, file=file.path(dataPath,(paste0("inputsMS/Afot1pVL_env_sort_", Sys.Date(),".csv"))))
```


