---
title: "MAKE START DATA for MS - R Notebook"
output: html_notebook
---

#INPUTS
Takes the datasets used in the workshop, and applies the learning that we need to start with:
- data below 100m only
- excluding Tromsøflaket (first cruise) data which has some analytical anomalies for the moment
- generate a random one subsample per VL dataset for first analysis step

## Libraries
```{r}
library(tidyverse)
library(oce)
```

# LOAD DATA

```{r}
#NB 2022-10-19 latest (all sample/Oslo) species cleaning applied so smaller spp dataset

env_orig <- read.csv(file.path(dataPath,"inputsMS/env_MS_2023-08-04.csv")) %>% as.data.frame
otu_orig <- read.csv(file.path(dataPath,"inputsMS/otu_MS_2023-08-04.csv")) %>% as.data.frame

#left join and split to ensure ordered identically
JoinedDat<- left_join(otu_orig,env_orig)
plot(JoinedDat$X, JoinedDat$Y)

```

## GENERATE 1 per VL DATASET

### SUBSAMPLE
```{r}
# Subset samples for maneagbility - Take a random (sub)sample from each VL

subSamp1perVL <- JoinedDat %>% group_by(VL) %>% slice_sample(n=1)
length(unique(subSamp1perVL$VL))

plot(subSamp1perVL$X, subSamp1perVL$Y)
```


### SPLIT INTO OTU & ENV DATASETS

```{r}
Afot1pVL_otu_sort<-subSamp1perVL %>% ungroup %>% select(c(2:which(colnames(subSamp1perVL)=="Weberella_bursa")))
Afot1pVL_env_sort<-subSamp1perVL %>% select(c(2,(which(colnames(subSamp1perVL)=="Weberella_bursa")+1):which(colnames(subSamp1perVL)=="sand")))  

```


write the subsample to csv with sys.Date appended to keep track of which is the active sample that day
```{r}
write.csv(Afot1pVL_otu_sort, file=file.path(dataPath,(paste0("inputsMS/otu_1pVL_MS_", Sys.Date(),".csv"))))
write.csv(Afot1pVL_env_sort, file=file.path(dataPath,(paste0("inputsMS/env_1pVL_MS_", Sys.Date(),".csv"))))
```


